Reaction-Advection-Diffusion Models of Air Pollution

### 2024-04-26 paper.tex
4

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{tikz}
\title{Reaction-Advection-Diffusion Models of Air Pollution}
\author{Daniel Henricks}

@@

\section{A Simple Model}

@@

We need to represent the sinks using some loss rate. Let's assume that the losses are proportional to the amount of the species in the box, so the losses are
equal to $km$. (It turns out trivially that $k = \frac{1}{\tau}$). Let's also assume that the sources are independent of $m$. Then, we yield the ODE

@@

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{tikz}
\title{Reaction-Advection-Diffusion Models of Air Pollution}
\author{Daniel Henricks}

@@

\subsection{A Simple Model}

@@

We need to represent the sinks using some loss rate. Let's assume that the losses are proportional to the amount of the species in the box, so the losses are
equal to $km$. (It turns out trivially that $k = \frac{1}{\tau}$). Let's also assume that the sources are independent of $m$ and are 
equal to some constant $S$. Then, we yield the ODE

@@

This steady state assumption is reasonable because the rate of change of the mass, $\frac{dm}{dt}$, tends to be very small in
comparison to the production rate. Because of this, we classify this steady state as a quasi steady state.

\subsection{A More Realistic Model: the Puff Model}

We will now investigate a puff model; such a model describes fluid elements moving in space. We define a fluid element
to be a volume of air in which all particles move with the same velocity. The most common application of a puff model is diffusion from a 
smokestack, something that is essential for understanding how air pollution diffusion. We define the mass balance equation for this model
to be

\begin{equation}
    \frac{dX_{c}}{dt} = E + P - L - D
\end{equation}

where $X_{c}$ is the concentration of element X. Assume that the puff takes a column shape and that the height of the column is $h$.
Also assume that there is no chemical production ($P$) and that the losses are proportional to the concentration at some rate $L = kX_{c}$. 
We then form the mass-balance equation:

\begin{equation}
    \frac{dX_{c}}{dt} = \frac{E}{h} - kX_{c}
\end{equation}
This is another separable ODE which is trivial to solve. The puff model's steady states are only valid in low-wind conditions, however (Zannetti, 1985).
Many computer systems used by the federal government to predict air pollution use variations of this model with more complicated parameters.

\section{A first diffusion model}

We will first review the diffusion model established by Tirabassi (1989). Analytical models such as this one use solutions to the 
advection-diffusion equation below:

\begin{equation}
    u\frac{\partial C}{\partial x} = \frac{\partial}{\partial y} (K_y \frac{\partial C}{\partial y}) + \frac{\partial}{\partial z} (K_z \frac{\partial C}{\partial z}) + S
\end{equation}

where the parameters are:

\begin{itemize}
    \item Mean velocity ($u$)
    \item Concentration ($C$)
    \item Source term ($S$)
    \item Lateral eddy exchange constant ($K_y$)
    \item Vertical eddy exchange constant ($K_z$)
 \end{itemize}

\section{InMAP: A more computationally efficient model for finding emissions}

One of the recent developments in the study of air pollution was the creation of computationally efficient models for computing the harm to human health from
diffusion. A prevalent model that has been implemented in industry by companies such as WattTime is InMAP \url{(http://spatialmodel.com/)}. 
InMAP is a model that evaluates pollution by simplifying typical CTMs via using an annual-average basis for its results.
